Theories of deliberated choice

### 2022-11-24 simple.tex
More attempts of Answ

An argumentative theory claims that whenever some of its arguments fail to convince $i$ of choosing $\pga$, it can argue something else.
It is said to be truthful iff this claim is correct.
More precisely, its claims are about paths of the sort $\hist = (\arga, …, \ar_2, …, \ar_3, …)$ where $\ar_2 \in \dgi(\arga)$, $\ar_3 \in \dgi(\ar_2)$, …
The following definition makes this precise by mandating that whenever increasing subsequences of $\hist$, involving arguments $\ar_j$ proposed by the theory, are not convincing, then the theory is able to produce yet another argument in defense of its claim.%
\footnote{Intuition.
If $a_1$, used by the theory, is not decisive, then the theory has an answer, namely, $a_2$, which works at least temporarily.
If $a_1$ used by th is not dec, and $a_2$ used by th is not dec *in the same sequence*, then $a_3$ needs to defend *in the same sequence*.
$(…, a_1)$ NO and $(…, a_1, …, a_2)$ NO then $a_3$ defends it.
$\alpha$ of the form $(\alpha^1, \alpha^2)$ where $a_1$ in end of $α^1$ NO and $a_2$ in end of $α^2$ NO.}
\begin{definition}[Truthfulness]
	\label{def:truth}
	An argumentative theory $(\pga, \arga, \dg)$ for the exposure protocol $(I, \allargs, {\leadsto})$ is \emph{truthful} iff 
	$\forall i \in I, \hist \in \allhist, n \in \N, k \in \intvl{1, \card \hist}^n, \ar \in \allargs^n$:
	\begin{equation}
		\left.\begin{aligned}
			&\ar_1 = \arga \\
			&\forall 1 ≤ j < n: \ar_{j + 1} = \dg(\ar_j, α_{\intvl{1, k_j}}) \in \histbend{k_{j + 1}} \\
			&\forall 1 ≤ j ≤ n: \hist_{\intvl{1, k_j}} \nileadsto \pga
		\end{aligned}\right\}
		⇒ \dg(\ar_n, \hist) ≠ \emptyset.
	\end{equation}
\end{definition}
Given the paths $\alpha, \alpha'$, the sequence $(\alpha, \alpha')$ denotes a path constituted of the arguments in $\alpha$ followed by the arguments of $\alpha'$.
\begin{axiom}[Answerability]
	\label{ax:ans}
	$\forall i \in I, \hist \in \allhist, \ar \in \gargsi$:
	\begin{equation}
		\left.\begin{aligned}
			&\hist \nileadsto \pga \\
			&(\hist, \ar) \ileadsto \pga \\
			&\forall \ar' \in \allargs: (\hist, \ar, \ar') \ileadsto \pga
		\end{aligned}\right\}
		⇒ \ar \ileadstost \pga.
	\end{equation}
\end{axiom}
\footnote{$a_1, a_2, a_3$ NOT and $a_1, a_2, a_3, a_4$ YES and $a_1', a_2, a_3, a_4$ NO imply that $a_1, a_2, a_3, a_4, a'_1$ NO.
Generalizing: $\alpha$ NO and $\alpha, a_4$ YES and $\alpha', a_4$ NO imply that $\exists a' \in \alpha' \suchthat \alpha, a_4, a'$ NO.
Let’s try with this too powerful axiom, but then we should weaken it, as it could be that $a_1$ AND $a_4$ TOGETHER are decisive, whence the axiom fails. We should have then some argument $a_1 + a_4$ to replace them, and…?
In summary, we want to avoid that $a$ is not decisive but we can’t observe it when starting with some $α$ even though $α$ has also no decisive arguments: if $α$ has no decisive argument, then it must be possible to counter it.

@@

@@

A path can be shortened to its first $k \in \intvl{0, \card{\hist}}$ elements, which is denoted by $\restr{\hist}{\intvl{1, k}}$, and “shortening” a path beyond its length yields the same path again, thus given $k \in \N$, $\restr{\hist}{\intvl{1, k}} = \restr{\hist}{\intvl{1, \min\set{\card{\hist}, k}}}$.
Define a path as being decisive for $p$, $\hist \ileadstost p$, iff $\hist \ileadsto p \land \forall \ar \in \allargs: (\hist, \ar) \ileadsto p$, or equivalently, iff $\forall \hist' \suchthat \card{\hist} ≤ \card{\hist'} ≤ \card{\hist} + 1 \land \restr{\hist'}{\intvl{1, \card{\hist}}} = \hist: \hist' \ileadsto p$. Thus, $\hist \nileadstost p ⇔ \exists \ar \in \allargs \suchthat (\hist, \ar) \nileadsto p$.

@@

\begin{theorem}[Validity]
	If an argumentative theory $(\pga, \arga, \dg)$ is truthful and satisfy \cref{ax:ans,ax:r} and $\allargs$ is finite, then the theory is correct.
\end{theorem}
\begin{proof}
	Consider any $i \in I$.
	We have to show that $\pga \in \Phi_i$.
	Assume depth $3$.
	Assume that $\arga$ and $\ar_2$ are not decisive. Let’s show that $\ar_3$ is decisive.
	Assume that $(a_4, \arga, \ar_2, a) \nileadsto \pga$.
	By TF, $(a_4, \arga, \ar_2, a, \ar_3, ?) \ileadsto \pga$ (where $?$ stands for any argument or nothing).
	By Reinst, $(a_4, \arga, \ar_3, ?) \ileadsto \pga$.
	By Self-s, $(a_4, \ar_3, ?) \ileadsto \pga$.

	Now consider $(\ar_5, \ar_2) \ileadsto \pga$. (Or any such path not ending with $\ar_3$.)
	By Answ, $\exists a \suchthat (\ar_5, \ar_2, a) \nileadsto \pga$.
	We obtain $(\ar_5, \ar_3) \ileadsto \pga$.
\end{proof}
\begin{proof}[Sketch of proof]
	Consider any $i$. We have to show that $\pga \in \Phi_i$.%
	\footnote{Assume $(a_4, \arga, a) \nileadsto \pga$ (thus, $a$ attacks $\arga$). 
	By tf, $(a_4, \arga, a, \ar_2, …) \ileadsto \pga$.
	We get $(a_3, \arga, a, \ar_2) \ileadsto \pga$ by answ.
	By reinst: $(\ar_2)$ YES.}

>>>>

An argumentative theory claims that whenever some of its arguments fail to convince $i$ of choosing $\pga$, it can argue something else.
It is said to be truthful iff this claim is correct.
More precisely, its claims are about paths of the sort $\hist = (\arga, …, \ar_2, …, \ar_3, …)$ where $\ar_2 \in \dgi(\arga)$, $\ar_3 \in \dgi(\ar_2)$, …
The following definition makes this precise by mandating that whenever increasing subsequences of $\hist$, involving arguments $\ar_j$ proposed by the theory, are not convincing, then the theory is able to produce yet another argument in defense of its claim.%
\footnote{Intuition.
If $a_1$, used by the theory, is not decisive, then the theory has an answer, namely, $a_2$, which works at least temporarily.
If $a_1$ used by th is not dec, and $a_2$ used by th is not dec *in the same sequence*, then $a_3$ needs to defend *in the same sequence*.
$(…, a_1)$ NO and $(…, a_1, …, a_2)$ NO then $a_3$ defends it.
$\alpha$ of the form $(\alpha^1, \alpha^2)$ where $a_1$ in end of $α^1$ NO and $a_2$ in end of $α^2$ NO.}
\begin{definition}[Truthfulness]
	\label{def:truth}
	An argumentative theory $(\pga, \arga, \dg)$ for the exposure protocol $(I, \allargs, {\leadsto})$ is \emph{truthful} iff 
	$\forall i \in I, \hist \in \allhist, n \in \N, k \in \intvl{1, \card \hist}^n, \ar \in \allargs^n$:
	\begin{equation}
		\left.\begin{aligned}
			&\ar_1 = \arga \\
			&\forall 1 ≤ j < n: \ar_{j + 1} = \dg(\ar_j, α_{\intvl{1, k_j}}) \in \histbend{k_{j + 1}} \\
			&\forall 1 ≤ j ≤ n: \restr{\hist}{\intvl{1, k_j}} \nileadsto \pga
		\end{aligned}\right\}
		⇒ \dg(\ar_n, \hist) ≠ \emptyset.
	\end{equation}
\end{definition}
Given the paths $\alpha, \alpha'$, the sequence $(\alpha, \alpha')$ denotes a path constituted of the arguments in $\alpha$ followed by the arguments of $\alpha'$.
\begin{axiom}[Answerability-1]
	$\forall …$:
	\begin{equation}
		\left.\begin{aligned}
			&(\ar_4, \ar_2, \ar) \nileadsto \pga \\
			&(\ar_4, \ar_2, \ar, \ar_3) \ileadstost \pga \\
			&(\ar_5, \ar_2, \ar) \nileadsto \pga
		\end{aligned}\right\}
		⇒ (\ar_5, \ar_2, \ar, \ar_3) \ileadstost \pga.
	\end{equation}
\end{axiom}
Requires excluding:
	\begin{equation}
		\left.\begin{aligned}
			&(\ar_4, \ar_2, \ar) \nileadsto \pga \\
			&(\ar_4, \ar_2, \ar, \ar_3, \ar') \ileadsto \pga \\
			&(\ar_5, \ar_2, \ar) \nileadsto \pga\\
			&(\ar_5, \ar_2, \ar, \ar_3, \ar') \nileadsto \pga.
		\end{aligned}\right\}
	\end{equation}

To exclude:
	\begin{equation}
		\left.\begin{aligned}
			&(\ar_4, \ar_2, \ar) \nileadsto \pga \\
			&(\ar_4, \ar_2, \ar, \ar_3) \ileadstost \pga \\
			&(\ar_5, \ar_2, \ar) \ileadsto \pga\\
			&(\ar_5, \ar_2, \ar_3, \ar') \nileadsto \pga
		\end{aligned}\right\}
	\end{equation}
Using:
	\begin{equation}
		\left.\begin{aligned}
			&(\ar_4, \ar_2, \ar) \nileadsto \pga \\
			&(\ar_4, \ar_2, \ar, \ar_3) \ileadstost \pga \\
			&(\ar_5, \ar_2, \ar) \ileadsto \pga
		\end{aligned}\right\}
		⇒ (\ar_5, \ar_2, \ar_3) \ileadstost \pga.
	\end{equation}

\begin{axiom}[Answerability]
	\label{ax:ans}
	$\forall i \in I, \hist \in \allhist, \ar \in \gargsi$:
	\begin{equation}
		\left.\begin{aligned}
			&\hist \nileadsto \pga \\
			&(\hist, \ar) \ileadsto \pga \\
			&\forall \ar' \in \allargs: (\hist, \ar, \ar') \ileadsto \pga
		\end{aligned}\right\}
		⇒ \ar \ileadstost \pga.
	\end{equation}
\end{axiom}
\footnote{$a_1, a_2, a_3$ NOT and $a_1, a_2, a_3, a_4$ YES and $a_1', a_2, a_3, a_4$ NO imply that $a_1, a_2, a_3, a_4, a'_1$ NO.
Generalizing: $\alpha$ NO and $\alpha, a_4$ YES and $\alpha', a_4$ NO imply that $\exists a' \in \alpha' \suchthat \alpha, a_4, a'$ NO.
Let’s try with this too powerful axiom, but then we should weaken it, as it could be that $a_1$ AND $a_4$ TOGETHER are decisive, whence the axiom fails. We should have then some argument $a_1 + a_4$ to replace them, and…?
In summary, we want to avoid that $a$ is not decisive but we can’t observe it when starting with some $α$ even though $α$ has also no decisive arguments: if $α$ has no decisive argument, then it must be possible to counter it.

@@

\begin{axiom}[Answerability - using newly built arguments (newest so far)]
	\label{ax:ansn}
	$\forall i \in I, \hist^j \in (\allhist)^n$:
	\begin{equation}
		[\forall j: \hist^j \ileadstost \pga] ⇒ [\exists \ar \in \allargs \suchthat \forall j: (\hist^j, \ar) \ileadstost \pga].
	\end{equation}
\end{axiom}

@@

A path can be shortened to its first $k \in \intvl{0, \card{\hist}}$ elements, which is denoted by $\restr{\hist}{\intvl{1, k}}$, and “shortening” a path beyond its length yields the same path again, thus given $k \in \N$, $\restr{\hist}{\intvl{1, k}} = \restr{\hist}{\intvl{1, \min\set{\card{\hist}, k}}}$.
Define a path as being decisive for $p$, $\hist \ileadstost p$, iff $\hist \ileadsto p \land \forall \ar \in \allargs: (\hist, \ar) \ileadsto p$, or equivalently, iff $\forall \hist' \suchthat \card{\hist} ≤ \card{\hist'} ≤ \card{\hist} + 1 \land \restr{\hist'}{\intvl{1, \card{\hist}}} = \hist: \hist' \ileadsto p$. Thus, $\hist \nileadstost p ⇔ \exists \ar \in \allargs \suchthat (\hist, \ar) \nileadsto p$ (or…?).

@@

\begin{theorem}[Validity]
	If an argumentative theory $(\pga, \arga, \dg)$ is truthful and satisfy \cref{ax:ans,ax:r} and $\allargs$ is finite, then the theory is correct.
\end{theorem}
\begin{proof}
	Consider any $i \in I$.
	We have to show that $\pga \in \Phi_i$.
	Assume depth $3$.
	Assume that $\arga$ and $\ar_2$ are not decisive. Let’s show that $\ar_3$ is decisive.
	Assume that $(a_4, \arga, \ar_2, a) \nileadsto \pga$.
	By TF, $(a_4, \arga, \ar_2, a, \ar_3) \ileadstost \pga$.
	By Reinst, $(a_4, \arga, \ar_2, \ar_3) \ileadstost \pga$.
	By Self-s, $(a_4, \arga, \ar_3) \ileadstost \pga$.
	By Self-s, $(a_4, \ar_3) \ileadstost \pga$.

	Now consider $(\ar_5, \ar_2) \ileadsto \pga$. (Or any such path not ending with $\ar_3$.)
	We want $(\ar_5, \ar_2, \ar_3) \ileadstost \pga$.
	By Answ, $\exists a \suchthat (\ar_5, \ar_2, a) \nileadsto \pga$.
\end{proof}
\begin{proof}[Sketch of proof]
	Consider any $i$. We have to show that $\pga \in \Phi_i$.%
	\footnote{Assume $(a_4, \arga, a) \nileadsto \pga$ (thus, $a$ attacks $\arga$). 
	By tf, $(a_4, \arga, a, \ar_2, …) \ileadsto \pga$.
	We get $(a_3, \arga, a, \ar_2) \ileadsto \pga$ by answ.
	By reinst: $(\ar_2)$ YES.}
